### 2020-12-03 Paper/main.tex
Fix typos and stuff

\documentclass[conf]{new-aiaa}
%\documentclass[journal]{new-aiaa} for journal papers
\usepackage[utf8]{inputenc}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\usepackage{longtable,tabularx}
\usepackage{fixme}
\setlength\LTleft{0pt}
\hypersetup{
  colorlinks=true
}
\usepackage{tikz}

\newcommand{\average}[1]{\langle #1 \rangle}

% Setup fixme env for collaborative notes
\fxsetup{status=draft, layout=inline, theme=color}

\title{Computation and comparison of the stable Northeastern US marine boundary layer}

\author{Lawrence C. Cheung\footnote{Principal member of technical staff, Thermal/Fluid Science \& Engineering, AIAA member}}
\affil{Sandia National Laboratories, Livermore, CA 94550}

\author{
  Shreyas Ananthan\footnote{Researcher V, High-Performance Algorithms and Complex Fluids, AIAA Member},
  Michael J. Brazell\footnote{Researcher III, High-Performance Algorithms and Complex Fluids, AIAA Member},
  Ganesh Vijayakumar\footnote{Researcher III, Mechanical Engineering, AIAA Member}, and
}
\affil{National Renewable Energy Laboratory, Golden CO 80401}

\author{Nathaniel B. deVelder\footnote{Postdoctoral appointee, Wind
    Energy Technologies Department, AIAA member} and Alan
  S. Hsieh.\footnote{Senior member of technical staff, Wind Energy
    Technologies Department, AIAA member}} \affil{Sandia National
  Laboratories, Albuquerque, NM 87185}

\begin{document}

\maketitle

\begin{abstract}
  In this work we investigate the behavior of stable marine boundary
  layers located near the coast of the Northeastern United States.
  Using the ExaWind LES codes, three stable atmospheric conditions
  were chosen to match the Cape Wind measurements of Archer \emph{et
    al.} with wind speeds of 5 m/s, 10 m/s, and 15 m/s at the 20m
  measurement height.  The behavior of the stable boundary layers,
  including mean flow quantities and turbulent statistics, are
  examined and compared to previous computations of the neutral and
  unstable offshore boundary layer at the same location.  This study
  also examines the domain and mesh requirements necessary to capture
  the turbulent scales for the stable offshore boundary layers.
  Finally, we compare solutions computed using both Nalu-Wind and
  AMR-Wind solvers, and report on their relative accuracy and
  performance in this study.

\end{abstract}

\section{Nomenclature}
\fxnote{Update symbols}
{\renewcommand\arraystretch{1.0}
  \noindent\begin{longtable*}{@{}l @{\quad=\quad} l@{}}
ABL       & Atmospheric Boundary Layer \\
$\alpha$  & Wind shear exponent  \\
$f$       & Frequency \\
$f_{max}$  & Maximum resolvable frequency due to mesh limits \\
$f_{Ny}$   & Nyquist frequency \\
$L$       & Turbulent integral length scale \\
$L_{Ob}$   & Obukhov length scale \\
LES       & Large Eddy Simulation \\
$u_\tau$   & Friction velocity    \\
$R_{ij}$   & Two point correlation tensor \\
SBL       & Stable Boundary Layer \\
$TI$      & Turbulence intensity \\
$TKE$     & Turbulent kinetic energy
\end{longtable*}}

\section{Introduction}
%\fxnote{Update this introduction.}

\lettrine{T}he planned installation of several offshore wind energy
plants in the United States has highlighted the need to properly
understand the wind resource and atmospheric characteristics of the US
Atlantic Coast.  Several atmospheric phenomena particular to this
region, such as coastal low-level jets or seasonal Nor’easters, have
the potential to substantially impact the operation and power
production of offshore wind plants.  Of particular interest to the
current study is the atmospheric stability of the marine boundary
layer in the Northeastern US.

Atmospheric stability plays a large role in determining the power
production of wind plants because it directly affects the vertical
distribution of momentum and turbulent energy in the atmospheric
boundary layer (ABL).  The differences between the stable, neutral, or
unstable stratified ABL can lead to large changes in wind speed or
turbulence profiles, and ultimately change the operation of wind
turbines.  Atmospheric stability may also play a role in the formation
of low-level jets \cite{nunalee2014mesoscale} and cause increased
fatigue loads on offshore wind turbines.

Several recent measurement campaigns have provided data to understand
the ABL and wind characteristics for potential offshore wind farms in
the Northeastern US.  Pichugina et al. \cite{pichugina2017properties}
measured the wind profiles and vertical shear profiles in the Gulf of
Maine using a ship-borne lidar approach.  Analysis of measured data
sets from Nantucket Sound by Archer et
al. \cite{archer2016predominance} showed a predominance of low-shear,
unstable conditions at the site.  However, strong seasonal variations
and stratification changes due to diurnal variation were also
observed.

\subsection{Previous Simulation Work}

In addition to these measurement campaigns, large eddy simulations (LES) have
also been used to study ABL stability characteristics.  Recent work by
Kaul et al. \cite{kaul2020large} has shown that LES computations using
Nalu-Wind can capture the neutral and convectively unstable onshore
ABL.  Previous simulations by Cheung et al. \cite{cheung2020large}
successfully replicated the unstable and neutral ABL corresponding to
the Cape Wind meteorological tower measurements
\cite{archer2016predominance} and showed the effects of surface
heating on atmospheric stratification. LES of the stable ABL (SBL) presents
certain challenges compared to neutral and convective stability states, stemming
primarily from small dominant eddy length scales and low turbulence intensity.
These challenges require higher resolution grids and/or advanced subgrid
scale (SGS) models. As such, LES research of the SBL has been slower
to evolve than the other stability states.

The first successful LES of the SBL was published by Mason and Derbyshire in
1990 \cite{Mason1990}. They used a strategy of simulating a near steady-state
neutral stability turbulent simulation, then added a cooling-surface boundary
to develop the stable ABL. They recorded a need for implausibly strong initial
perturbations to sustain turbulence using the Smagorinsky SGS model. Brown et
al.\ \cite{Brown1994} worked with Mason and Derbyshire to add stochastic backscatter
to the SGS model, resulting in better agreement with surface layer observations.
Andren \cite{Andren1995}, simulated two SBL types, one with a neutral layer aloft, and
one capped by an inversion, comparing two SGS models. He found significant improvements
with the Sullivan SGS model, on par with the earlier backscatter model. Kemp and
Thompson \cite{Kemp1996} used the same model as \cite{Brown1994} to study particle
dispersion, but the LES code did not replicate an expected plume growth. Derbyshire
wrote a review of SBL modeling along with future projections in 1999 \cite{Derbyshire1999},
noting the importance of local Richardson number and identifying a need to go
to higher resolutions in SBL research.

In 2000, Kosovic and Curry \cite{Kosovic2000} used an LES simulation with a nonlinear SGS
model containing backscatter to favorably compare to the \fxnote{[Nate: define BASE]} BASE measurements, showing that
the SBL could reach a quasi steady-state. They also showed good agreement with the Nieustadt
analytical model. In the same year, Saiki et al. \cite{Saiki2000} simulated an SBL with
a strong capping inversion at z=500m, producing reasonable surface layer profiles and
identifying a need for SBL specific SGS models. Beare and MacVean \cite{Beare2004}
presented the first grid-focused work on LES of the SBL, using grids with deltas
between 2m and 12.5m, finding significant variation between grid sizes. The inclusion
of a backscatter model resulted in significant variation at a 6.25m resolution, but
very little at a 2m resolution. Basu et al. \cite{Basu2005} published a set of
turbulence statistics, measurements, and LES simulations in support
of the viability of Nieustadt's local scaling hypothesis under very stable
conditions. Jimenez and Cuxart \cite{Jimenez2005} set out to quantify the range
for which standard SGS models, using the Kolmogorov theory for dissipation, were
valid for increasingly stable flows. They found the model to perform adequately
for weakly stable, and some moderately stable flow regimes.

An inter-comparison of LES simulations of the SBL was organized in 2006, the
results of which are published in Beare et al.\ \cite{Beare2006}. Collectively,
this is known as the GABLS initiative, and involved simulating a moderately
stable flow taken from Kosovic and Curry \cite{Kosovic2000}. Eleven codes
were compared in all, wherein grid deltas of less than 12.5m were found to
successfully maintain resolved turbulence, grid deltas less than 3.125m
showed little statistical change, and results became independent of SGS
modeling at 1m resolution. In the same year, Kumar et al. \cite{Kumar2006}
successfully simulated an entire diurnal cycle with LES using a Lagrangian
dynamic scale dependent SGS model. This model has been used by several
researchers since 2011, including Lu and Porte-Agel \cite{Lu2011}, Huang
and Bou-Zeid \cite{Huang2013}, Park et al.\ \cite{Park2014}, and
Abkar and Porte-Agel \cite{Abkar2015}. Stoll and Porte-Agel \cite{Stoll2007}
compared three grid refinements and three scale dependent dynamic SGS models,
showing that the Lagrangian and locally averaged models reproduced important
flow features. The work of van Dop and Axelsen \cite{vanDop2007} provided
even more evidence in support Nieuwstadt's local scaling hypothesis for SBLs.
Chow and Zhou \cite{Chow2011} showed evidence that the Smagorinsky model is
simply too dissipative for use in strongly stable flows.

More recent work in LES of the SBL includes Matheou \cite{Matheou2016} who
explored a large parameter space using the Smagorinsky-Lilly SGS
closure, as well as Sullivan et al.\ \cite{sullivan2016turbulent}. Several researchers
have applied SBL simulation in the wind energy space \cite{Aitken2014, Abkar2015, Ghaisas2017}.
In 2020, Wurps et al. \cite{Wurps2020} provided a rigorous grid resolution study for the SBL,
looking at mean flow, turbulence variables, and two-point correlations, using a dynamic
SGS model that allows backscatter. They found grid requirements to be 2.5m, 10m, and 20m
for stable, neutral, and convective boundary layers respectively. They advocate looking
at two point correlations in order to quantify flow structure size, which in turn
provides a basis for evaluating grid resolution. In all of the reviewed literature, SBL LES
has been run in the onshore context only, a complete comparison
including offshore stable ABL conditions has yet to be completed.

\subsection{Objectives}
In the current study on stable marine boundary layers, we aim to build
on the previous work in the literature and complete several
objectives.  First, a series of stable boundary layer conditions will
be computed using the Exawind solvers which match the measurements of
Archer et al. \cite{archer2016predominance}.  The behavior of the
stable boundary layers, including mean flow quantities and turbulent
statistics, are examined and compared to the neutral and unstable
cases computed in \cite{cheung2020large}.  Secondly, we examine the
domain and mesh requirements necessary to capture the turbulent scales
for the stable offshore boundary layers.  Finally, two LES solvers
within the Exawind project are compared in this study: the
unstructured, incompressible Navier-Stokes solver Nalu-Wind and the
structured, adaptive mesh refinement solver AMR-Wind.  Both codes are
used to calculate the same SBL, and the relative differences between
the codes are reported.

%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION 2: METHODOLOGY
%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{methodology}

%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION 3: RESULTS
%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{results}

\section{Conclusion}
In this study we have simulated the stable offshore boundary layer
matching the characteristics of measurements from Nantucket Sound in
the Northeastern US.  Three cases with wind speeds of 5 m/s, 10 m/s,
and 15 m/s were selected, and simulations were performed using the
ExaWind codes to match the TI and stability levels observed at from
the Cape Wind platform.  When examining the effect of stratification
using the Obukhov length, differences in the ABL behavior were seen
bewteen the very stable boundary layer cases (5 m/s and 10 m/s), and
the stably stratified case (15 m/s).  The very stable cases had
shorter turbulent structures and less growth in the boundary layer
compared to the 15 m/s case.  The behavior of the three stable
boundary layers was also compared to the neutral and unstable cases
studied in \cite{cheung2020large}.  As expected, the turbulent
structures in all stable cases were noticeably smaller than in the
neutral and unstable boundary layers.

In addition to these simulations, we also examined the domain and mesh
requirements necessary for the stable offshore boundary layers
considered here.  Based on the calculated turbulent integral
lengthscale, horizontal domain sizes of 750m $\times$ 750m were found
to be necessary for these simulations.  Various mesh resolutions were
also considered for the stable runs, and it was determined that the
finest resolution of $dx$=2.5m was able to resolve the longitudinal
length scales and highest temporal frequencies, but inadequate for
resolving the turbulence in the lateral direction.  Further study is
warranted to comprehensively determine the impact of stable
stratification on the required mesh resolution.

Lastly, a comparison of the ExaWind solvers Nalu-Wind and AMR-Wind was
conducted using the stable offshore boundary layer at 5 m/s.  The mean
statistics were found to agree well between the two LES codes.  Small
differences were found in wind profiles at higher elevations and for
the peak $S_u$ spectra, and excellent matches were observed for the
averaged veer, temperature, and TI profiles.

%% \section*{Appendix}

%% An Appendix, if needed, should appear before the acknowledgments.

\section*{Acknowledgments}

The authors wish to acknowledge the contributions from M. Churchfield
and S. Yellapantula for their development of the LES boundary
conditions and for their efforts in validating the ExaWind codes.

This research was supported by the Wind Energy Technologies Office of
the US Department of Energy Office of Energy Efficiency and Renewable
Energy.  Sandia National Laboratories is a multimission laboratory
managed and operated by National Technology \& Engineering Solutions
of Sandia, LLC, a wholly owned subsidiary of Honeywell International
Inc., for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-NA0003525. The views expressed in the
article do not necessarily represent the views of the U.S. Department
of Energy or the United States Government.

This work was authored in part by the National Renewable Energy
Laboratory, operated by Alliance for Sustainable Energy, LLC, for the
U.S. Department of Energy (DOE) under Contract No. DE-AC36-
08GO28308. M. Brazell was funded by the Exascale Computing
Project (17-SC-20-SC), a collaborative effort of two DOE organizations
(Office of Science and the National Nuclear Security Administration).

A portion of the research was performed using computational
resources sponsored by the Department of Energy's Office of Energy
Efficiency and Renewable Energy and located at the National Renewable
Energy Laboratory.

\bibliography{references}

\end{document}